Staged Abstract Interpreters

### 2019-04-04 paper/60_cases.tex
paper revision

\paragraph{Original AC} \todo{WIP}

The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA
for a CPS language consisting of lambda terms, applications, @letrec@ and
primtive operators. The analyses for different syntactic constructs are
decomposed into different functions, such as @analyzeCall@ and @analyzeApp@. The
idea of closure generation is to rewrite these functions. Where previously they
may take both static arguments and dynamic arguments, after the rewrite only the
static arguments is taken. In this case, the static arguments are syntactic
terms; the dynamic arguments are stores. After written in AC style, functions
like @compCall@ (compiled version of @analyzeCall@) return a value of type
@CompAnalysis@, i.e., a closure that takes a store and returns a store. The
result of multiple calls on such functions, for example, @compCall@ and
@compArgs@, can be composed. The generated closure only takes stores, because
the input program is specialized into the closure. The code is shown in
Figure~\ref{compare_ac_sai} (left).

@@

On the other side, our approach does exactly the same thing through staging: the
syntactic terms are static, and stores are dynamic, therefore the generated code
just looks-up and updates the store. Figure~\ref{compare_ac_sai} (right) shows
the code written with LMS. In fact, the only changes are the type of @Store@ is
replaced with @Rep[Store]@ indicating that the values of type @Store@ will be
known at the next stage. Indeed, additional engineering efforts are required to
make this happen, including: the staged version @Map@ which is already included
in LMS; implicit @lift@ function that transform a current-stage constant value
to next stage; next-stage representation of the syntactic terms, i.e., proper
@toString@ functions of AST structs. We consider these efforts relatively small,
and they do not interfere the actual analysis we desire.

@@

The target language we presented in Section~\ref{bg_lang} is essentially a
higher-order functional language. One fundamental analysis task for functional
programs is control-flow analysis, i,e., determining which functions will
possibly be applied at each call-site. The abstract interpreter we used in
Section~\ref{unstaged_abs} and Section~\ref{sai} is a store-widened 0-CFA-like
abstraction, moreover it is also a pushdown control-flow analysis; in the last
section, we also reviewed AC with finite-state 0-CFA. In this section, based on
the existing staged abstract interpreter, we further develop the staging
techniques with control-flow analyses, including recoverying a more precise
store model and adding context-sensitivity to the analysis.

@@

\subsection{Modular Analysis for Free}

>>>>

\paragraph{Original AC}
The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA
for a small CPS language consisting of abstractions, applications, recursion and
primtive operators. Figure \ref{compare_ac_sai} (left) shows the AC
implementation: different syntactic constructs are handled in different functions
(e.g., @compCall@, @compApp@, etc.).
Before refactoring into AC, these functions would take the store as a
regular argument; now every function returns a value of type @CompAnalysis@,
which is a function (closure) value that takes and returns stores. After the
first run of analysis, we have traversed the AST and obtained a single closure
takes a store and returns a store. Therefore, the ASTs are static terms, and the
stores are dynamic terms. Now we can invoke the compiled closure with an initial
store to complete the analysis.

@@

On the other side, our approach does exactly the same thing through staging: the
AST are static, and stores are dynamic, therefore we annotate the type @Store@
with @Rep@ (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
changes of our appraoch using LMS framework are the types of @Store@ to
@Rep[Store]@, since they will be known at the next stage. The generated code
will just looks-up and update the store. The functions at current stage are
eliminated in the residual program, therefore we do not generate higher-order
functions, which provides additional performance improvement. Nevertheless,
the support from LMS framework is necessary, such as staging support for @Map@,
we consider these efforts are reusable and complete orthogonal to the
implementation of analysis.

@@

In this section, to demonstrate the SAI approach is widely applicable, we extend
and refactor the staged abstract interpreter we presented in Section \ref{sai}
to various control-flow analysis (CFA). We first discuss calling-context
sensitive analysis for staged analyzers; then we will see how to implement
store-widening with staging, which is a common optimization in CFA; finally, we
integrate abstract garbage collection with staging, \todo{where the addresses to
  be collected can be computed at the staging time}.

\subsubsection{Context-sensitivity}

\subsubsection{Store-widened Analysis}

\subsubsection{Abstract Garbage Collection at Staging Time}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

@@

\subsection{Modular Analysis for Free} \label{modular}
